---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
workingDirectory <- c("C:\\Users\\Chirag\\chirag personal\\US\\coursera\\RepData\\RepData_PeerAssessment1")
outputDirectory <- c("C:\\Users\\Chirag\\chirag personal\\US\\coursera\\RepData\\RepData_PeerAssessment1\\")

filename <- c("activity.csv")
setwd(workingDirectory)

stepData <- read.csv(file = filename, na.strings="NA")
stepData$interval <- factor(stepData$interval)



```
## What is mean total number of steps taken per day?

```{r}
tot_steps <- aggregate(steps ~ date, stepData, FUN=sum)

barplot(tot_steps$steps, names=tot_steps$date, col="blue", main="Total Daily Steps",xlab="Date", ylab="Total Steps")

#mean_steps <- format(mean(tot_steps$steps, na.rm=TRUE), digits=2, nsmall=2)

#median_steps <- median(tot_steps$steps, na.rm=TRUE)

```
### The Average number of steps taken daily (Mean) = `r format(mean(tot_steps$steps, na.rm=TRUE), digits=2, nsmall=2)`  
### The Median number of steps taken daily (Median) = `r median(tot_steps$steps, na.rm=TRUE)`

## What is the average daily activity pattern?
```{r}
dailyAvg <- aggregate(steps ~ interval, stepData, FUN = mean)
plot(dailyAvg$interval, dailyAvg$steps, type="n", main="Average Daily Activity Pattern",xlab="Interval", ylab="Average Steps")
lines(dailyAvg$interval, dailyAvg$steps, type="l")

maxAvg <- max.col(dailyAvg$steps)
```

### The Interval containing the maximum average daily steps is `r maxAvg`   

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
